Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-14 vesicle_code/docs/adhesion/adhesion.tex
started some small latex changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo[inline]{$\sigma$ is overdefined as the Hamacher constant, and the
tension, and possibly the stress.  Replace Hamacher constant with $H$.}

\todo[inline]{In the xtensional flow, put a plus sign after the 0.
Indicate the flow with arrows.  Do the same for the shear and Figure 9
(right).}

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been widely utilized
as biological cell mimics in biophysics and material engineering \cite{}. 
%
The ever expanding applications have encouraged detailed experimental, theoretical and numerical investigations of vesicle hydrodynamics
in applied flow, electric and magnetic fields. Theoretical investigations often take the form of small-deformation analysis or spheroidal dynamics of a nearly-spherical vesicle (or nearly-circular in two dimensions) subjected to an applied flow or under an external forcing \cite{Barthes-BieselRallison1981_JFM,Misbah2006_PRL,Vlahovska2007_PRE,Finken2008_EPL,ZhangZahnTanLin2013_PoF,Nganguia2013_PRE}. 
On the other hand, various numerical methods have been developed for simulating the transient hydrodynamics of vesicle suspensions  \cite{BagchiJohoson2005_JBE,Biben2005_EJP,Veerapaneni2009_JCP,SeolHuKimLai2016_JCP}.  
%Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively investigated. In a planar shear flow, the vesicle hydrodynamics is characterized by
excess area (reduced volume), viscosity contrast between interior and exterior fluids, and shear rate of the imposed far-field fluid flow. In addition a vesicle
with a rigid particle inside is also investigated as a biological mimic of a 
eukaryotic cell with a nucleus that occupies nearly half of the intracellular volume \cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast and shear rate. At high viscosity contrast the tank-tread (TT) dynamics transitions to tumbling (TB) \cite{Misbah2006_PRL,Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the vesicle suspension \cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations \cite{GhigliottiBibenMisbah2010_JFM} and experiments \cite{KantslerSegreSteinberg2008_EPL,ZabuskySegreDeschamps2011_PoF}.
Between TT and TB vesicle hydrodynamics, a breathing (tremble) is also observed \cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2011_JFM,SpannZhaoShaqfeh2014_PoF} to affect the effective shear viscosity in a 
different way.
In an extensional flow (planar or uniaxial), vesicle shape dynamics
depends sensitively on the vesicle reduced volume  and the elastic capillary number \cite{KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2013_JFM,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape and oscillatory undulation of the vesicle membrane are two examples of the complex vesicle hydrodynamics in an extensional flow.

Collective dynamics of vesicless naturally is dictated by the individual vesicle-vesicle interactions.
%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
In numerical simulations of vesicle suspensions, it is often assumed that the liquid film between vesicles acts to keep the
membrane-membrane interaction, such as the van der Waals type adhesive interaction, negligible.
However the amphiphilic natures of lipid molecules (a hydrophobic tail and a hydrophilic head with an electric dipole) complicate the interactions between a lipid bilayer membrane and 
another bilayer membrane \cite{EvansMetcalfe1984_BJ,Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces} or
a solid (such as a glass substrate or a nano particle).
Studies confirmed that such interaction potential inevitably consists of a long-range attraction component and a short-range repulsion component 
\cite{Book_IntermolecularSurfaceForces}, and the combination of the two gives rise to adhesion at the separation distance that minimizes the
interaction potential \cite{Book_IntermolecularSurfaceForces}. Such adhesive potentials are essential in many biomedical, biological and biophysical processes.
For example, vesicle adhesion is essential for membrane fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

%
Adhesion between a vesicle and a substrate has been extensively studied, primarily focused on analyzing the static equilibrium shapes \cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase} than the the transient adhesion process; exceptions include \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by Evans and Metcalfe \cite{EvansMetcalfe1984_BJ}, who measured 
the reduction in the free energy per unit area of membrane-membrane contact formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion energy per unit area to the bending energy (bending rigidity of lipid bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle deformation in the presence of adhesion \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In this work we focus on regimes where the ratio is of order unity, between the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$) regimes. 
In this regime the membrane deformation may increase the ``contact
area'' between two vesicles and enhance the adhesion effects. The
equilibration of vesicle membranes under adhesion in these regimes have
been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their Van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong Van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation distance
\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

Boundary integral equation (BIE) approaches are well-suited for solving the low Re flow problems considered here as they lead to reduction in dimensionality and 
achieve high-order accuracy even for moving geometry problems. When the vesicles adhere, one major issue for BIE solvers is to resolve the vesicle-vesicle hydrodynamic forces which become 
{\em nearly singular}. In this work, we use the recently developed spectrally-accurate algorithm of \cite{barnett2015spectrally} for overcoming this problem. To overcome the numerical stiffness induced by the membrane bending forces, 
we extend the adaptive time-stepping scheme developed in \cite{quaife2016adaptive} to our setting.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adhesion Model}
The adhesion model of Sukumaran and Seifert~\cite{suk-sei2001} considers adhesion
between a single three-dimensional vesicle and a solid wall at $z=0$.
\todo[inline]{Need a cartoon picture in here.  Some kind of version of
the one I had for my SIAM Life Sciences Talk}
The adhesion potential in their model is of the functional form in equation~\ref{eq:adhesion_potential} with $(m,n) = (4,2)$:
\begin{align}
\label{eq:adhesion_potential}
  \phi(z) = \sigma \left[ 
    \left(\frac{\delta}{z}\right)^m - \frac{m}{n} \left(\frac{\delta}{z}\right)^n \right],
\end{align}
where $z$ is the distance between the solid wall and the vesicle,
$\sigma$ is the Hamaker constant, and $\delta$ is the adhesion length
scale.  We generalize this model to an unbounded suspension of vesicles.
We assume that the adhesion force applies between all pairs of points on
different vesicles, so we are interested in Hamaker constants that are
appropriate for lipid-lipid interactions. In
Appendix~\ref{sec:appendixA}, the adhesive force at a point $\xx$ on
vesicle $j$ is shown to be
\begin{align}
  \AA\xx:=-\sigma m \delta^{n}\sum_{\substack{k=1 \\ k \neq j}}^M 
  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^{m+2}} 
  \left(\delta^{m-n} - \|\xx - \yy\|^{m-n} \right) ds_\yy.
  \label{eqn:adhesionForce}
\end{align}

%\begin{align}
%  \AA\xx:=-4 \sigma \delta^2\sum_{\substack{k=1 \\ k \neq j}}^M 
%  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^6} 
%  \left(\delta^2 - \|\xx - \yy\|^2 \right) ds_\yy.
%  \label{eqn:adhesionForce}
%\end{align}

The exponents $(m,n)$ depend on the geometry of the two objects under adhesion \cite{Book_IntermolecularSurfaceForces}: $(m,n)=(4,2)$ corresponds to two
flat, planar surfaces interacting with each under a long-range attraction and a short-range repulsion. On the other hand, $(m,n) = (12,6)$ corresponds to the L-J potential between
two molecules. The adhesion potential in  the integral of equation~\ref{eqn:adhesionForce} is between two small patches of lipid bilayer membrane. Thus we argue that
a reasonable choice for $(m,n)$ is between $(m,n) = (4,2)$ and $(m,n) = (12,6)$.

@@

When two vesicles move towards (or away from) each other under a constant force $F$ without any external fluid flow, the height  $(h)$ of the thin liquid film between two vesicles follows the
draining dynamics \cite{RamachandranLeal2010_PoF}
\begin{equation}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{equation}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of exterior fluid, $R_0$ is the radius of the undeformed vesicle.
For the case of a constant forcing $F$ (independent of separation distance $h$ and time $t$), equation~\ref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{equation}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{equation}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction between two vesicles,  and consequently $h(t)$ decreases (increases) from  the initial separation distance $h(0)$. 
When the force is attractive, $h$ decreases monotonically and $t$ in the above
equation is the
draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{equation}
\label{eq:Fh}
F(h) \sim \sigma \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{equation}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion is dominant at ``small" scales ($h<\delta$).
Integration of equation~\ref{eq:dhdt} with the above $F(h)$ gives the relationship between $t$ and $h$.
In this case the relationship involves an integral of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = -\frac{1}{\sigma^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align}
\begin{equation}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{\sigma^{1/3}\delta^{5/3}}.
\end{equation}
Equation~\ref{eq:teq_scaling} tells us that both adhesion strength $\sigma$ and the separation distance $\delta$ affect
the time it takes for a pair of vesicles to reach equilibrium under adhesion force $F$ in equation~\ref{eq:Fh}.  In particular, the equilibration time $t$ 
is proportional to (1) the separation distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion strength $\sigma^{-1/3}$ 
(as in constant forcing case \cite{RamachandranLeal2010_PoF}).  From the above analysis we also expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength is independent of the specific form of the potential.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling01.eps}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling02.eps}
\caption{Scaling of the time it takes for two vesicles to reach equilibrium: (a) scaling with respect to $\delta$ with $\sigma=0.1$, and (b) scaling
with respect to $\sigma$ with $\delta = 0.4$. Arrows indicate the time by when the equilibrium is reached within $1\%$.}
\label{fig:qflow00}
\end{figure}

To test the scaling of the equilibration time $t$ with respect to $\sigma$ and $\delta$, we simulate the vesicle adhesion dynamics in a quiescent flow.
Starting with two vesicles at a distance of twice the radius at $t=0$, the long-range attraction pulls the vesicles together. Figure~\ref{fig:qflow00}(a) plots
the scaled adhesion energy versus time with $\sigma=0.1$ and three values of $\delta$ as labeled. 
Adhesion energy reaches minimum at equilibrium, and the scaled $E_{adh}$ evolves towards zero at equilibrium. 
The arrow indicate the times when the scaled adhesion energy
reaches within $1\%$ of equilibrium: $t\sim 96$ for $\delta=0.2$, $t\sim 45$ for $\delta = 0.3$, and $t\sim 30$ for $\delta=0.4$:
\begin{align}
\frac{t_{\delta=0.2}}{t_{\delta=0.4}} \sim \frac{96}{30}=3.2 &\Longleftrightarrow \left(\frac{0.4}{0.2}\right)^{5/3}\sim 3.17,\\
\frac{t_{\delta=0.2}}{t_{\delta=0.3}} \sim \frac{96}{45}=2.1 &\Longleftrightarrow \left(\frac{0.3}{0.2}\right)^{5/3}\sim 2.
\end{align}

The scaling with respect to adhesion strength $\sigma$ is illustrated in figure~\ref{fig:qflow00}(b), where $\delta = 0.4$ and $\sigma$ varies from $0.1$ to $0.5$ as labeled.
Again the arrows indicate the times when the equilibrium is reached within $1\%$:
\begin{align}
\frac{t_{\sigma=0.1}}{t_{\sigma=0.5}} \sim \frac{34}{20} = 1.71 &\Longleftrightarrow \left(\frac{0.5}{0.1}\right)^{1/3}\sim 1.71.
\end{align}
From the above results we conclude that the scaling in equation~\ref{eq:teq_scaling} captures the adhesion dynamics of two vesicles that interact with each other via the adhesion force in equation~\ref{eq:Fh}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{In the extensional flow, put a plus sign after the 0.
Indicate the flow with arrows.  Do the same for the shear and Figure 9
(right).}

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been widely utilized
as biological cell mimics in biophysics and material engineering \cite{}. 
%
The ever expanding applications have encouraged detailed experimental,
theoretical and numerical investigations of vesicle hydrodynamics in
applied flow, electric and magnetic fields. Theoretical investigations
often take the form of small-deformation analysis or spheroidal dynamics
of a nearly-spherical vesicle (or nearly-circular in two dimensions)
subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM,Misbah2006_PRL,Vlahovska2007_PRE,Finken2008_EPL,ZhangZahnTanLin2013_PoF,Nganguia2013_PRE}.
On the other hand, various numerical methods have been developed for
simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE,Biben2005_EJP,Veerapaneni2009_JCP,SeolHuKimLai2016_JCP}.  
%Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL,Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,ZabuskySegreDeschamps2011_PoF}.
Between TT and TB vesicle hydrodynamics, a breathing (tremble) is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2011_JFM,SpannZhaoShaqfeh2014_PoF}
to affect the effective shear viscosity in a different way.  In an
extensional flow (planar or uniaxial), vesicle shape dynamics depends
sensitively on the vesicle reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL,ZhaoShaqfeh2013_JFM,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}:
Asymmetric shape and oscillatory undulation of the vesicle membrane are
two examples of the complex vesicle hydrodynamics in an extensional
flow.

Collective dynamics of vesicles naturally is dictated by the individual vesicle-vesicle interactions.
%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
In numerical simulations of vesicle suspensions, it is often assumed
that the liquid film between vesicles acts to keep the membrane-membrane
interaction, such as the van der Waals type adhesive interaction,
negligible.  However the amphiphilic natures of lipid molecules (a
hydrophobic tail and a hydrophilic head with an electric dipole)
complicate the interactions between a lipid bilayer membrane and another
bilayer
membrane~\cite{EvansMetcalfe1984_BJ,Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces}
or a solid (such as a glass substrate or a nano particle).  Studies
confirmed that such interaction potential inevitably consists of a
long-range attraction component and a short-range repulsion
component~\cite{Book_IntermolecularSurfaceForces}, and the combination
of the two gives rise to adhesion at the separation distance that
minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. Such adhesive
potentials are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is essential for membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.

%
Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their Van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong Van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

Boundary integral equation (BIE) approaches are well-suited for solving
the low Re flow problems considered here as they lead to reduction in
dimensionality and achieve high-order accuracy even for moving geometry
problems. When the vesicles adhere, one major issue for BIE solvers is
to resolve the vesicle-vesicle hydrodynamic forces which become {\em
nearly singular}. In this work, we use the recently developed
spectrally-accurate algorithm of~\cite{barnett2015spectrally} for
overcoming this problem. To overcome the numerical stiffness induced by
the membrane bending forces, we extend the adaptive time-stepping scheme
developed in~\cite{quaife2016adaptive} to our setting.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adhesion Model}
The adhesion model of Sukumaran and Seifert~\cite{suk-sei2001} considers adhesion
between a single three-dimensional vesicle and a solid wall at $z=0$.
\todo[inline]{Need a cartoon picture in here.  Some kind of version of
the one I had for my SIAM Life Sciences Talk} The adhesion potential in
their model is of the functional form in
equation~\eqref{eq:adhesion_potential} with $(m,n) = (4,2)$:
\begin{align}
\label{eq:adhesion_potential}
  \phi(z) = H \left[ 
    \left(\frac{\delta}{z}\right)^m - \frac{m}{n} \left(\frac{\delta}{z}\right)^n \right],
\end{align}
where $z$ is the distance between the solid wall and the vesicle,
$H$ is the Hamaker constant, and $\delta$ is the adhesion length
scale.  We generalize this model to an unbounded suspension of vesicles.
We assume that the adhesion force applies between all pairs of points on
different vesicles, so we are interested in Hamaker constants that are
appropriate for lipid-lipid interactions. In
Appendix~\ref{sec:appendixA}, the adhesive force at a point $\xx$ on
vesicle $j$ is shown to be
\begin{align}
  \AA\xx:=-H m \delta^{n}\sum_{\substack{k=1 \\ k \neq j}}^M 
  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^{m+2}} 
  \left(\delta^{m-n} - \|\xx - \yy\|^{m-n} \right) ds_\yy.
  \label{eqn:adhesionForce}
\end{align}

%\begin{align}
%  \AA\xx:=-4 H \delta^2\sum_{\substack{k=1 \\ k \neq j}}^M 
%  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^6} 
%  \left(\delta^2 - \|\xx - \yy\|^2 \right) ds_\yy.
%  \label{eqn:adhesionForce}
%\end{align}

The exponents $(m,n)$ depend on the geometry of the two objects under
adhesion~\cite{Book_IntermolecularSurfaceForces}: $(m,n)=(4,2)$
corresponds to two flat, planar surfaces interacting with each under a
long-range attraction and a short-range repulsion. On the other hand,
$(m,n) = (12,6)$ corresponds to the L-J potential between two molecules.
The adhesion potential in  the integral of
equation~\eqref{eqn:adhesionForce} is between two small patches of lipid
bilayer membrane. Thus we argue that a reasonable choice for $(m,n)$ is
between $(m,n) = (4,2)$ and $(m,n) = (12,6)$.

@@

When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height $(h)$ of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{equation}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{equation}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of
exterior fluid, $R_0$ is the radius of the undeformed vesicle.  For the
case of a constant forcing $F$ (independent of separation distance $h$
and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{equation}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{equation}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction
between two vesicles,  and consequently $h(t)$ decreases (increases)
from  the initial separation distance $h(0)$.  When the force is
attractive, $h$ decreases monotonically and $t$ in the above equation is
the draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{equation}
\label{eq:Fh}
F(h) \sim H \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{equation}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion
is dominant at ``small" scales ($h<\delta$).  Integration of
equation~\eqref{eq:dhdt} with the above $F(h)$ gives the relationship
between $t$ and $h$.  In this case the relationship involves an integral
of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3
  F^{1/3}} = -\frac{1}{H^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align}
\begin{equation}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{H^{1/3}\delta^{5/3}}.
\end{equation}
Equation~\eqref{eq:teq_scaling} tells us that both adhesion strength $H$
and the separation distance $\delta$ affect the time it takes for a pair
of vesicles to reach equilibrium under adhesion force $F$ in
equation~\eqref{eq:Fh}.  In particular, the equilibration time $t$ is
proportional to (1) the separation distance $\delta^{-5/3}$ (thus
$t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion
strength $H^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling01.eps}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling02.eps}
\caption{\label{fig:qflow00} Scaling of the time it takes for two
  vesicles to reach equilibrium. {\em Left}: Scaling with respect to
  $\delta$ with $H=0.1$.  {\em Right}: Scaling with respect to $H$ with
  $\delta = 0.4$. Arrows indicate the time by when the equilibrium is
  reached within $1\%$.}
\end{figure}

To test the scaling of the equilibration time $t$ with respect to $H$
and $\delta$, we simulate the vesicle adhesion dynamics in a quiescent
flow.  Starting with two vesicles at a distance of twice the radius at
$t=0$, the long-range attraction pulls the vesicles together.
Figure~\ref{fig:qflow00}(a) plots the scaled adhesion energy versus time
with $H=0.1$ and three values of $\delta$ as labeled.  Adhesion energy
reaches minimum at equilibrium, and the scaled $E_{adh}$ evolves towards
zero at equilibrium.  The arrow indicate the times when the scaled
adhesion energy reaches within $1\%$ of equilibrium: $t\sim 96$ for
$\delta=0.2$, $t\sim 45$ for $\delta = 0.3$, and $t\sim 30$ for
$\delta=0.4$:
\begin{align}
\frac{t_{\delta=0.2}}{t_{\delta=0.4}} \sim \frac{96}{30}=3.2 &\Longleftrightarrow \left(\frac{0.4}{0.2}\right)^{5/3}\sim 3.17,\\
\frac{t_{\delta=0.2}}{t_{\delta=0.3}} \sim \frac{96}{45}=2.1 &\Longleftrightarrow \left(\frac{0.3}{0.2}\right)^{5/3}\sim 2.
\end{align}

The scaling with respect to adhesion strength $H$ is illustrated in
Figure~\ref{fig:qflow00}(b), where $\delta = 0.4$ and $H$ varies from
$0.1$ to $0.5$ as labeled.  Again the arrows indicate the times when the
equilibrium is reached within $1\%$:
\begin{align}
\frac{t_{H=0.1}}{t_{H=0.5}} \sim \frac{34}{20} = 1.71 &\Longleftrightarrow \left(\frac{0.5}{0.1}\right)^{1/3}\sim 1.71.
\end{align}
From the above results we conclude that the scaling in
equation~\eqref{eq:teq_scaling} captures the adhesion dynamics of two
vesicles that interact with each other via the adhesion force in
equation~\eqref{eq:Fh}.

\subsection{Effects of the adhesion parameters on equilibrium configuration of a vesicle pair}
\label{subsec:qflow_adhesion_parameters} 
In the simulations of a pair of 2d vesicles interacting with each other
via the adhesion force in equation~\eqref{eq:Fh}, two vesicles are
initially placed not too far apart  from each other for numerical
convenience.  Without any external forcing such as an imposed electric
field or a fluid flow in the far-field, the long-range attraction pulls
two vesicles together until their separation distance is close to
$\delta$ (where $F$ vanishes).  Depending on the adhesion strength
relative to the membrane bending rigidity, the vesicle may deform
significantly in the contact region while still maintain nearly
spherical shape on the opposite side to the contact
region~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,
Book_IntermolecularSurfaceForces,
RamachandranAndersonLealIsraelachvili2010_Langmuir}.
\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
\includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_01.jpeg}
\caption{Two vesicles of length $\pi$ with $\kappa=1$, $H=5$ and
  $\delta=0.2$.  (a) Equilibrium shapes of two vesicles under adhesive
  interactions with adhesion strength $H=0.5$ and separation distance
  $\delta=0.2$.  Vesicle reduced area $\Delta A=0.95$, $0.75$ and $0.6$
  as labeled.  The color coding is the tension along the vesicle. (b)
  Membrane shapes in the contact region.}
\label{fig:Dec18_vesicle_shape}
\end{figure}

Figure~\ref{fig:Dec18_vesicle_shape}(a) shows the equilibrium shapes of
vesicles with three values of reduced area: $\Delta A=0.95$, $0.75$ and
$0.6$ with $H=5$ \todo[inline]{Is this really the Hamacher constant?
It's not 0.5?} and $\delta = 0.2$.  The equilibrium vesicle shape for
$\Delta A=0.95$ is the circular cap with a flatten contact region.  This
is similar to the observed shapes of two vesicles under strong adhesive
interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
